Interactive Visual Analysis of Lumbar Back Pain \subtitle{What the Lumbar Spine Tells About Your Life}

### 2014-10-01 draft.tex
Replaced Associations

% \abstract{The abstract should summarize the contents of the paper and should contain at least 70 and at most 200 words.}
\abstract{ 
Epidemiology aims to provide insight into disease causations.
%%
Hence, subject groups (cohorts) are analyzed to correlate the subjects' varying lifestyles, their medical properties and diseases.
%%
Recently, these cohort studies comprise medical image data.
%%
We assess potential relations between image-derived variables of the lumbar spine with lower back pain in a cross-sectional study.
%%
Therefore, an Interactive Visual Analysis (IVA) framework was created and tested with 2,540 segmented lumbar spine data sets.
%%
The segmentation results are evaluated and quantified by employing shape-describing variables, such as the spine canal course, its curvature and torsion.
%%
We analyze mutual dependencies among shape-describing variables and non-image variables, e.g. pain indicators.
%%
Therefore, we automatically extract a decision tree classifier for each non-image variable.
%%
We provide an IVA technique to compare classifiers with interactive scatter plot visualizations.
%%
As a first result, we conclude that image-based variables are only sufficient for classification of lifestyle factors within the data.
%%
A correlation between lumbar spine shape and lower back pain was not existent in the automatically trained classifiers.
%%
However, the presented approach is a valuable extension for the IVA of epidemiological data.
%%
Hence, associations between non-image variables were successfully detected and characterized.
}

@@

We extract associations between spine shape and back pain characteristics.

@@

	\item Statistical methods, such as regression analysis, assess the association of selected variables with the investigated disease.

@@

Modern cohort studies often comprise medical image data.
%%
The analysis of these data is challenging.
%These data are hard to analyze.
%%
Individual diagnosis or manual segmentation of each body structure by radiologists is tedious, costly and comprises little reproducibility.
%%
Segmentation algorithms are not generally available and need to be carefully adapted for each body structure.
%%
Segmentation data is usually analyzed by abstracting it into key figures, such as diameters or distances.
%%
These numeric values can be compared with non-image variables to retrieve correlations.
%%--------------------------------
\subsection{Lower Back Pain}
%%
%The lower (\emph{lumbar}) spine is the most stressed spine part .
%%
Lower back pain is one of the most frequent complaints in the Western civilization \cite{Hoy2010}.
%%
Epidemiologists assume associations between lumbar back pain and lifestyle factors.
%%
These include \emph{nutrition}, \emph{physical activity} and \emph{body posture combined with physical stress at work}.
%%
The exact causes as well as particularly vulnerable risk groups are unknown.
%%
%Potential \emph{confounding} effects are also subject of current research.
%%

@@

We calculated an association matrix to assess correlations between the image variables.

>>>>

% \abstract{The abstract should summarize the contents of the paper and should contain at least 70 and at most 200 words.}
\abstract{ 
Epidemiology aims to provide insight into disease causations.
%%
Hence, subject groups (cohorts) are analyzed to correlate the subjects' varying lifestyles, their medical properties and diseases.
%%
Recently, these cohort studies comprise medical image data.
%%
We assess potential relations between image-derived variables of the lumbar spine with lower back pain in a cross-sectional study.
%%
Therefore, an Interactive Visual Analysis (IVA) framework was created and tested with 2,540 segmented lumbar spine data sets.
%%
The segmentation results are evaluated and quantified by employing shape-describing variables, such as the spine canal course, its curvature and torsion.
%%
We analyze mutual dependencies among shape-describing variables and non-image variables, e.g. pain indicators.
%%
Therefore, we automatically extract a decision tree classifier for each non-image variable.
%%
We provide an IVA technique to compare classifiers with interactive scatter plot visualizations.
%%
As a first result, we conclude that image-based variables are only sufficient for classification of lifestyle factors within the data.
%%
A correlation between lumbar spine shape and lower back pain was not existent in the automatically trained classifiers.
%%
However, the presented approach is a valuable extension for the IVA of epidemiological data.
%%
Hence, relations between non-image variables were successfully detected and characterized.
}

@@

We extract possible associations between spine shape and back pain characteristics.

@@

	\item Statistical methods, such as regression analysis, assess the correlation of selected variables with the investigated disease.

@@

Modern cohort studies often comprise medical image data.
%%
The analysis of these data is challenging.
%These data are hard to analyze.
%%
Individual diagnosis or manual segmentation of each body structure by radiologists is tedious, costly and comprises little reproducibility.
%%
Segmentation algorithms are not generally available and need to be carefully adapted for each body structure.
%%
Segmentation data is usually analyzed by abstracting it into key figures, such as diameters or distances.
%%
These numeric values can be compared with non-image variables to retrieve correlations.
%%--------------------------------
\subsection{Lower Back Pain}
%%
%The lower (\emph{lumbar}) spine is the most stressed spine part .
%%
Lower back pain is one of the most frequent complaints in the Western civilization \cite{Hoy2010}.
%%
Epidemiologists assume correlations between lumbar back pain and lifestyle factors.
%%
These include \emph{nutrition}, \emph{physical activity} and \emph{body posture combined with physical stress at work}.
%%
The exact causes as well as particularly vulnerable risk groups are unknown.
%%
%Potential \emph{confounding} effects are also subject of current research.
%%

@@

We calculated an correlation matrix to assess correlations between the image variables.
